---
title: "4_report_script_3"
author: "Paul Cuchot"
date: "02/05/2023"
output: github_document
# output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Test model with simulated data


### Library

```{r message=FALSE, warning=FALSE}
# bayesian modeling + plot
require(R2jags)
require(MASS)
require(mcmcplots)

# general
require(tidyverse)
require(bayesplot)

```

### Simulate data

#### Assumption for simulated data 

- Same number of breeders between years (5 pairs)
- Mean number of eggs = 8 (sd = 5)
- All chicks fledge and survive
- Adults survive during the breeding season
- Between years,  sessions are on the same days 
- Birds do not migrate or immigrate during the breeding season
- Same capture probability between juveniles and adults 

The function used to simulate data is described at the end of this document. 

```{r}
# load function 
source("2_function_sim_data.R")
# number of year to simlate
Nyears = 10

prod <- simul_data(
  
  # 5 pairs of breeders per year
  n_breeders = 5,
  # Nyears
  n_years = Nyears,
  # mean laying date for this site 
  mean_ld_site = 120,
  
  # CES start (julian days)
  start_ces = 100,
  # CES end (julian days)
  end_ces = 220,
  # sessions per year 
  n_session = 5)
```


### plot productivity through time (sim data)

Productivity throught breeding period for a single site. The vertical lines represent the simulated mean laying for each year. 

```{r pressure, echo=FALSE}

prod$capt_sess%>%
  ggplot(aes(x = t, y = prod, color = as.character(year)))+
  geom_point()+
  # add laying dates
  geom_vline(data = prod$mean_ld_year,
             aes(xintercept = mean_ld, 
                 color = as.character(year)), alpha = 0.8)+
  geom_line(alpha = 0.3)+
  scale_color_discrete(name = "SIM years")+
  xlab("Time (in Julian days)")+ylab("Productivity (N juveniles / N tot) ")+
  theme_light()

```

### Structure data for JAGS

```{r}

prod_f <- prod$capt_sess %>%
  mutate(an = as.numeric(year))

# data for the model
data <- list(nt = prod_f$n_capt_juveniles+prod_f$n_capt_adults,
             n0 = prod_f$n_capt_juveniles,
             date = as.numeric(prod_f$t),
             N = nrow(prod_f),
             N_an = length(unique(prod_f$an)),
             an = prod_f$an
)
```


### Define model 
```{r}
model <- "model{
 
  # loop on capture session
 
  for(i in 1:N){
    
    ## likelihood
    n0[i] ~ dbin(p[i], nt[i])
    
    p[i] <- asig[an[i]]/(1+exp((csig[an[i]]-date[i])/dsig[an[i]]))
  }
 
  # loop on an

  for (ii in 1:N_an){
    
    # asig[ii] <- alpha_asig + random_asig_an[an_n[ii]]
    # csig[ii] <- alpha_csig + random_csig_an[an_n[ii]]
    # dsig[ii] <- alpha_dsig + random_dsig_an[an_n[ii]]
    
    asig[ii] <- alpha_asig + random_asig_an[ii]
    csig[ii] <- alpha_csig + random_csig_an[ii]
    dsig[ii] <- alpha_dsig + random_dsig_an[ii]
    
    random_asig_an[ii] ~ dnorm(0, tau_asig_an)
    random_csig_an[ii] ~ dnorm(0, tau_csig_an)
    random_dsig_an[ii] ~ dnorm(0, tau_dsig_an)
 
  }

  alpha_asig ~ dnorm(0.7,0.1)T(0,1)
  alpha_csig ~ dnorm(150,0.1)T(100,200)
  alpha_dsig ~ dnorm(7,0.1)T(0,15)
 
#folded Cauchy version
  sigma_asig_an ~ dt(0, 0.01, 1)
  tau_asig_an <- pow(sigma_asig_an, -2)
 
#folded Cauchy version
  sigma_csig_an ~ dt(0, 0.01, 1)
  tau_csig_an <- pow(sigma_csig_an, -2)
  
  sigma_dsig_an ~ dt(0, 0.01, 1)
  tau_dsig_an <- pow(sigma_dsig_an, -2)

}
"
```

### Create initial values and parameters to save

```{r}
init_f <-  function(){
  
  list(
    
    alpha_asig = runif(1, 0, 1),
    alpha_dsig = runif(1, 3, 10),
    alpha_csig = rnorm(1, 150, 20),
    
    sigma_asig_an = rnorm(1, 3, 20),
    sigma_csig_an = rnorm(1, 3, 20),
    sigma_dsig_an = rnorm(1, 3, 20),
    
    random_asig_an = rnorm(Nyears, 0, 0.001),
    random_csig_an = rnorm(Nyears, 0, 0.001),
    random_dsig_an = rnorm(Nyears, 0, 0.001)
    
  )
}

inits <- list(init1 = init_f(), init2=init_f(), init3=init_f())

# parameter to save 
parameters3 <- c("asig","csig","dsig",
                 "random_asig_an","random_csig_an","random_dsig_an",
                 "sigma_csig_an","sigma_dsig_an","sigma_asig_an")

```

### Run model

```{r}
# run model 
md_1 <- jags(data = data,
             parameters.to.save = parameters3,
             model.file = textConnection(model),
             inits = inits,
             n.chains = 3,
             n.iter = 10000,
             n.burnin = 3000)
```

### Look at convergence 

```{r}
md_1
```



### Quickly compare mean xmid with 'real pheno'

```{r}
data.frame(estim = md_1$BUGSoutput$mean$c, 
           real = prod$mean_ld_year$mean_ld, 
           year = 1:Nyears)%>%
  ggplot(aes(x = estim, y = real, label = year))+
  geom_text(vjust = 1.5)+
  geom_point()+
  theme_bw()+
  stat_smooth(method = lm)
```


## Annexe

### Function to simulate data

```{r eval=F, echo=T}
simul_data <- function(n_breeders,
                       n_years, 
                       n_session, 
                       start_ces = 80,
                       end_ces = 120,
                       mean_ld_site = 120){
  
  # mean laying date (among breeding individuals - change between years)
  mean_ld <- round(rnorm(n_years, mean_ld_site, 15)) # (real pheno)
  
  sd_ld <- 7
  
  # number of eggs per pair
  mean_eggs <- 8
  sd_eggs <- 5
  
  # final data_set
  df_site <- data.frame(t = NA,
                        n_capt_adults = NA,
                        n_capt_juveniles = NA, 
                        prod = NA,
                        year = NA)[0,] 
  
  for(k in 1:n_years){
    
    # sample n_breeders laying events
    ld_dates <- round(rnorm(n_breeders, 
                            mean = mean_ld[k], 
                            sd = sd_ld))
    
    # fledglings dates 
    fledgl_dates <- round(ld_dates)+35
    
    # number of eggs per pair
    n_eggs <- abs(round(rnorm(n_breeders, mean_eggs, sd_eggs)))
    
    # create a dataframe (one row per breeding pair)
    df_breed <- data.frame(
      ld_dates = ld_dates,
      n_eggs = n_eggs,
      fledgl_dates = fledgl_dates
    )
    
    # sample (as CES design) 
    
    # choose days for capture session
    t_capt <- round(seq(start_ces, end_ces, 
                        length.out = n_session))
    
    # Number of sample individual per session (~capture effort)
    mean_n_capt <- 9
    
    # Dataframe with n_adults and n_juveniles captured per session
    df_session <- data.frame(t = t_capt,
                             n_capt_adults = NA,
                             n_capt_juveniles = NA,
                             prod = NA,
                             year = as.character(k)) 
    
    
    for(i in t_capt){
      
      # catchable adults
      n_adults <- n_breeders * 2
      
      # catchable juveniles
      n_juveniles <- sum(df_breed[df_breed$fledgl_dates < i ,]$n_eggs)
      
      # sample birds among available individuals
      capt_indiv <- sample( # how many individuals captured
        c(rep(0,n_adults), 
          rep(1,n_juveniles)),
        round(rnorm(1,mean_n_capt)), replace = TRUE)
      
      # how many adults
      df_session[df_session$t ==i,"n_capt_adults"] <- sum(capt_indiv == 0)
      # how many juveniles
      df_session[df_session$t ==i,"n_capt_juveniles"] <- sum(capt_indiv == 1) 
      
    }
    
    # calculate productivity
    df_session <- df_session%>%
      mutate(prod = n_capt_juveniles/(n_capt_adults+n_capt_juveniles))
    
    df_site <- rbind(df_site, df_session)
  }
  
  # df with mean ld per year ("real breeding date")
  df_mean_ld <- data.frame(year = 1:n_years,
                           mean_ld = mean_ld)
  
  return(list(capt_sess = df_site, 
              mean_ld_year = df_mean_ld))
  
}
```


